---
title: "Phydro testing PML functions"
author: "Jaideep Joshi"
date: "2023-05-27"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(rphydro)
library(tidyverse)
library(ggplot2)
library(reshape2)
```

```{r}
kphio = 0.087;        # quantum yield efficiency
ppfd = 1000;          # umol/m2/s
vpd  = 1000;         # Pa
co2  = 400;          # ppm
elv  = 0;            # m.a.s.l.
fapar = 0.7;         # fraction
rdark = 0.015;
psi_soil = 0;

par_cost = list(alpha=0.1, gamma=1);
par_plant = list(conductivity=3e-17, psi50=-2, b=2);

options = list(gs_method = "GS_IGF", 
               et_method = "ET_DIFFUSION",
               ftemp_vj_method = "FV_kumarathunge19",
               ftemp_rd_method = "FR_heskel16",
               ftemp_br_method = "FB_atkin15")

```

## Test meterological functions

### Latent heat of vaporization of water

```{r}
tibble(tc = -5:55) %>% mutate(dat = purrr::map_dbl(.x=tc, .f = ~calc_enthalpy_vap(.x))) %>% ggplot(aes(x=tc, y=dat/1e6)) + geom_line() + labs(y="lv [MJ kg-1]", x="T [degC]")
```



### Saturation vapour pressure

```{r}
tibble(tc = -5:100) %>% mutate(dat = purrr::map_dbl(.x=tc, .f = ~calc_esat(.x, 1.01325e5))) %>% ggplot(aes(x=tc, y=dat/1e3)) + geom_line() + labs(y="Esat [kPa]", x="T [degC]")
```



### Density of dry and moist air

```{r}
tibble(tc = -5:55) %>% 
  mutate(dat1 = purrr::map_dbl(.x=tc, .f = ~calc_density_air(.x, 1.01325e5, 5e3, T))) %>% 
  mutate(dat2 = purrr::map_dbl(.x=tc, .f = ~calc_density_air(.x, 1.01325e5, 0, T))) %>% 
  mutate(dat3 = purrr::map_dbl(.x=tc, .f = ~calc_density_air(.x, 1.01325e5, Inf, F))) %>% 
  ggplot(aes(x=tc)) + 
    geom_line(aes(y=dat1), col="brown") + 
    geom_line(aes(y=dat2), col="blue") + 
    geom_line(aes(y=dat3), col="yellow3") + 
    labs(y="Density [kg m-3]", x="T [degC]")
```

### Specific heat capacity of moist air

```{r}
tibble(tc = -5:100) %>% 
  mutate(dat1 = purrr::map_dbl(.x=tc, .f = ~calc_cp_moist_air(.x))) %>%
  ggplot(aes(x=tc)) + 
    geom_line(aes(y=dat1*1e-3), col="black") + 
    labs(y="Sp. heat cap [kJ kg-1 K-1]", x="T [degC]")
```


### Slope of saturation pressure vs temp curve

```{r}
tibble(tc = -5:55) %>% 
  mutate(dat1 = purrr::map_dbl(.x=tc, .f = ~calc_sat_slope(.x))) %>%
  ggplot(aes(x=tc)) + 
    geom_line(aes(y=dat1*1e-3), col="black") + 
    labs(y="s [kPa K-1]", x="T [degC]")
```



### Slope of saturation pressure vs temp curve

```{r}
tibble(tc = -5:55) %>% 
  mutate(dat1 = purrr::map_dbl(.x=tc, .f = ~calc_sat_slope(.x))) %>%
  ggplot(aes(x=tc)) + 
    geom_line(aes(y=dat1*1e-3), col="black") + 
    labs(y="s [kPa K-1]", x="T [degC]")
```


### Psychrometric constant

```{r}
tibble(tc = -5:55) %>% 
  mutate(dat1 = purrr::map_dbl(.x=tc, .f = ~calc_psychro(.x, 1.01325e5))) %>%
  ggplot(aes(x=tc)) + 
    geom_line(aes(y=dat1), col="black") + 
    labs(y="s [Pa K-1]", x="T [degC]")
```


### Aerodynamic conductance

```{r}
list(h_canopy = seq(1,50,length.out=10), v_wind = 0:10) %>% cross_df() %>% 
  mutate(dat1 = purrr::map2_dbl(.x=h_canopy, .y=v_wind, .f = ~calc_g_aero(.x, .y, .x+10))) %>%
  ggplot(aes(x=v_wind, y=dat1*1.01e5/8.3/300, group=h_canopy, col=h_canopy)) + 
    geom_line() + 
    labs(y="g_aero [mol m-2 s-1]", x="Wind speed [m s-1]")
```


### PML Transpiration from gs

```{r}
list(ga = exp(seq(log(1), log(800), length.out=20)), gs=1.6*exp(seq(log(0.01), log(0.4), length.out=20))) %>% cross_df() %>% 
  mutate(gs_ms = gs*8.31*298/1.01e5) %>% 
  mutate(ga_ms = ga*8.31*298/1.01e5) %>% 
  mutate(dat1 = purrr::map2_dbl(.x=ga_ms, .y=gs_ms, .f = ~calc_transpiration_pml(ppfd/2, .y, .x, 25, 1.01e5, vpd))) %>%
  ggplot(aes(x=gs, y=dat1/calc_enthalpy_vap(25)*55.5, group=ga, col=ga)) + 
    geom_line() + 
    geom_line(aes(y=gs*vpd/1.01e5), col="red") + 
    scale_colour_gradient(trans = "log") + 
    labs(y="T [mol m-2 s-1]")
```



### PML Transpiration from gs when vpd = 0

```{r}
list(ga = exp(seq(log(1), log(800), length.out=20)), gs=1.6*exp(seq(log(0.01), log(0.4), length.out=20))) %>% cross_df() %>% 
  mutate(gs_ms = gs*8.31*298/1.01e5) %>% 
  mutate(ga_ms = ga*8.31*298/1.01e5) %>% 
  mutate(dat1 = purrr::map2_dbl(.x=ga_ms, .y=gs_ms, .f = ~calc_transpiration_pml(ppfd/2, .y, .x, 25, 1.01e5, 0))) %>%
  ggplot(aes(x=gs, y=dat1/calc_enthalpy_vap(25)*55.5, group=ga, col=ga)) + 
    geom_line() + 
    geom_line(aes(y=gs*0/1.01e5), col="red") + 
    scale_colour_gradient(trans = "log") + 
    labs(y="T [mol m-2 s-1]")
```



### PML Transpiration from gs when ppfd = 0

```{r}
list(ga = exp(seq(log(1), log(800), length.out=20)), gs=1.6*exp(seq(log(0.01), log(0.4), length.out=20))) %>% cross_df() %>% 
  mutate(gs_ms = gs*8.31*298/1.01e5) %>% 
  mutate(ga_ms = ga*8.31*298/1.01e5) %>% 
  mutate(dat1 = purrr::map2_dbl(.x=ga_ms, .y=gs_ms, .f = ~calc_transpiration_pml(0, .y, .x, 25, 1.01e5, 1e3))) %>%
  ggplot(aes(x=gs, y=dat1/calc_enthalpy_vap(25)*55.5, group=ga, col=ga)) + 
    geom_line() + 
    geom_line(aes(y=gs*vpd/1.01e5), col="red") + 
    scale_colour_gradient(trans = "log") + 
    labs(y="T [mol m-2 s-1]")
```